---
title: "Use the MTPS Package"
author: "Li Xing, Xiaowen Cao, Shuai You, Mary L Lesperance, Xuekui Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MTPS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

Stacking is an ensemble machine learning algorithm, which allows
multiple prediction tasks to borrow information from each other. The Standard Stacking (SS) for multiple outcomes prediction problem consists of two steps of model fitting. In step 1, the individual models are fitted to predict resistance of each drug separately. In step 2, the combiner models are fitted to integrate information from predictions of individual models and to make final predictions. Two variations of stacking algorithm are proposed. They are Cross-Validation Stacking (CVS) and Residual Stacking (RS). 

The motivation for Cross-Validation Stacking (CVS) is that the combiner models are learned from the Step one ‘fitted’ values $\hat Y$ in SS, and the final predictions are calculated from the step one ‘predicted’ values using new data. The two types of quantities are different, for instance, in linear regression models, it is known that predicted values have larger variance than fitted values.  In CVS, The ‘fitted’ values $\hat Y$ in SS step 1 is replaced the cross-validation predicted values $\hat Y ^ *$ to deal with the discrepancy between fitted values and predicted values.

Sometimes in SS, the step 1 prediction accuracy may be compromised by
stacking when step 1 already provides accurate predictions. To
ameliorate this problem, Residual Stacking (RS) is proposed. In RS, all
information learned from step 1 models is retained. In step 2, the
outcome is learned using all other fitted value exclude the outcome as
the predictors to predict the residual of the outcome. Then, the
predictions in step 1 will be revised using the residuals.

The MTPS package is a flexible package which implements the
Revised Stacking algorithms. It can fit models and make predictions
using the above two variations of stacking algorithms and their
combination. It provides various base learners to train models in step
1 and step 2. Users can modify default parameters in base learners
provided in the package. It is also available for users to create new
base learners and use those learners to fit models. This package can fit models on continuous, binary and mix outcomes. An interface for model evaluation using cross-validation method is available in the package.

We extend the MTPS(version: 1.0.1) for predicting multiple survival
outcomes in May 2023. We update the original R package, MTPS, to a newer version(1.1.1), and now it can handle three types of outcomes: continuous, binary and survival outcomes.

# Example Data

As with any package, you first need to load it with the following command

```{r library, echo = T, results = 'hide', warning=FALSE, message=FALSE}
library(MTPS)
```
```{r library2, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
library(dplyr)
library(purrr)
library(tidyverse)
library(caret)
library(ggplot2)
library(reshape2)
```
The HIV Drug Resistance Database is used as the example data in the
package. In the HIV database, the resistance of five Nucleoside RT
Inhibitor (NRTI) drugs were used as multivariate outcomes, including
Lamivudine (3TC), Abacavir(ABC), Zidovudine (AZT), Stavudine (D4T),
Didanosine (DDI). The mutation variables are used as the predictors.
Some mutation variables were removed as they do not contain enough
variation. The final outcome data is a matrix of size 1246 × 5, and the predictor data is a matrix of 1246 × 228 values, which is provided in the package called "HIV". In the example data in the package, "YY" refers the outcome data and "XX" refers the predictor data.

```{r data}
data("HIV")
head(YY)
XX[8:9, 7:10]
dim(YY)
dim(XX)
```

# Model Fittig and Prediction

## Revised Stacking Algorithm for Continuous Outcome

The HIV data set is used as the example. To illustrate how to fit the
model and make predictions, we first split the HIV data into 2 parts,
the training data and testing data.

```{r ctn-split-data, echo = T, results = 'hide'}
set.seed(12345)
xmat <- XX %>% as.matrix()
ymat <- YY %>% as.matrix()
nobs <- nrow(xmat)
id <- XX %>% 
  rowMeans() %>% 
  createFolds(k = 5, list = FALSE)
training_id <- seq_len(nobs) %>% 
  sample(size = 0.8 * nobs)

y_train <- ymat[training_id, ]
y_test  <- ymat[-training_id, ]
x_train <- xmat[training_id, ]
x_test  <- xmat[-training_id, ]
```

The ` r multiFit` function fits individual models for each
outcome separately, referring the non-stacking algorithm. The following
code fits generalized linear models with elastic net regularization on
each outcome.

```{r ctn-noss, eval=FALSE, results = 'hide', eval=T}
 # no stacking
 fit_mult <- multiFit(xmat = x_train, ymat = y_train, method = glmnet1, family = "gaussian")

```

To set up the stacking algorithm, A list of algorithms for step one and step two need to be specified. For continuous outcome we need to specify ` r family = "gaussian"`. We also use ` r cv` and ` r residual` argument to specify whether we want to use Cross-Validation Stacking (CVS) or Residual Stacking (RS) or their combination. The default value for ` r cv` is "FALSE" and for ` r residual` is "TRUE", referring to the residual stacking
algorithm. The following code fits models using Standard Stacking
algorithm (SS), Cross-Validation Stacking (CVS), Residual Stacking (RS)
and Cross-Validation Residual Stacking (CVRS) with ` r MTPS` function. 

In the example, generalized linear models with elastic net
regularization are used to fit models in the step one and tree models
are applied in the step two for each outcome.

```{r ctn-train, eval=T, echo=T, message=FALSE, warning=FALSE, results='hide'}
# Standard Stacking 
fit_ss <- MTPS(xmat = x_train, ymat = y_train, family = "gaussian",
                            cv = FALSE, residual = FALSE,
                            method_step1 = glmnet1,
                            method_step2 = rpart1)
# Cross-Validation Stacking 
fit_cv <- MTPS(xmat = x_train, ymat = y_train, family = "gaussian",
                            cv = TRUE, residual = FALSE,
                            method_step1 = glmnet1,
                            method_step2 = rpart1)
# Residual Stacking 
fit_rs <- MTPS(xmat = x_train, ymat = y_train, family = "gaussian",
                            cv = FALSE, residual = TRUE,
                            method_step1 = glmnet1,
                            method_step2 = rpart1)
# Cross-Validation Residual Stacking 
fit_cvrs <- MTPS(xmat = x_train, ymat = y_train, family = "gaussian",
                            cv = TRUE, residual = TRUE,
                            method_step1 = glmnet1,
                            method_step2 = rpart1)
```

```{r echo=F,eval=T}
data("Internal")
```

The ` r predict` function returns the predicted value for the outcome on the testing data.

```{r ctn-predict, echo = T, results = 'hide', eval=T}
# no stacking
pred_mult <- predict(fit_mult, x_test)
# Standard Stacking 
pred_ss <- predict(fit_ss, x_test)
# Cross-Validation Stacking 
pred_cv <- predict(fit_cv, x_test)
# Residual Stacking 
pred_rs <- predict(fit_rs, x_test)
# Cross-Validation Residual Stacking 
pred_cvrs <- predict(fit_cvrs, x_test)
```

To visualize the above prediction, we plot the predicted value versus the actual outcomes of testing data for above algorithms.

```{r ctn-outcome,  echo = T, eval=T, message=FALSE, warning=FALSE}
n_test <- nrow(x_test)

ctn_plot_data_matrix <- cbind(
  rbind(pred_mult, pred_ss, pred_cv, pred_rs, pred_cvrs),
  y_test[rep(seq_len(n_test), 5), ]
)
ctn_plot_data <- data.frame(
  method = rep(c("No-Stacking", "SS", "CV", "RS", "CVRS"), each = n_test),
  ctn_plot_data_matrix
)
colnames(ctn_plot_data) <- c("method", paste0("pred.", colnames(y_test)), colnames(y_test))

dm1 <- ctn_plot_data %>%
  select(method, ABC, `3TC`, AZT, D4T, DDI) %>%
  pivot_longer(cols = -method, names_to = "Y", values_to = "yVal")
dm2 <- ctn_plot_data %>%
  select(method, pred.ABC, pred.3TC, pred.AZT, pred.D4T, pred.DDI) %>%
  pivot_longer(cols = -method, names_to = "Y", values_to = "predictVal")

ctn_plot_data <- bind_cols(dm1, select(dm2, -method))
colnames(ctn_plot_data) <- c("method", "Y", "yVal", "predict", "predictVal")
ctn_plot_data <- ctn_plot_data %>%
  mutate(
    method = factor(method, levels = unique(method)),
    yVal = as.numeric(yVal),
    predictVal = as.numeric(predictVal)
  )

ggplot(ctn_plot_data) +
  geom_point(aes(x = predictVal, y = yVal, color = method), size = 0.5, alpha = 0.8) +
  geom_abline(slope = 1, alpha = 0.2) +
  coord_fixed() +
  ylab("Testing Data Outcome") + xlab("Predicted Outcome on Testing Data") +
  scale_x_discrete(breaks = NULL) +
  scale_y_discrete(breaks = NULL) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank()
  ) +
  facet_grid(Y ~ method)
```


```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE, results='hide'}
n_test <- nrow(x_test)

ctn_plot_data_matrix <- cbind(
  rbind(pred_mult, pred_ss, pred_cv, pred_rs, pred_cvrs),
  y_test[rep(seq_len(n_test), 5), ]
)
ctn_plot_data <- data.frame(
  method = rep(c("No-Stacking", "SS", "CV", "RS", "CVRS"), each = n_test),
  ctn_plot_data_matrix
)
colnames(ctn_plot_data) <- c("method", paste0("pred.", colnames(y_test)), colnames(y_test))

dm1 <- ctn_plot_data %>%
  select(method, ABC, `3TC`, AZT, D4T, DDI) %>%
  pivot_longer(cols = -method, names_to = "Y", values_to = "yVal")
dm2 <- ctn_plot_data %>%
  select(method, pred.ABC, pred.3TC, pred.AZT, pred.D4T, pred.DDI) %>%
  pivot_longer(cols = -method, names_to = "Y", values_to = "predictVal")

ctn_plot_data <- bind_cols(dm1, select(dm2, -method))
colnames(ctn_plot_data) <- c("method", "Y", "yVal", "predict", "predictVal")
ctn_plot_data <- ctn_plot_data %>%
  mutate(
    method = factor(method, levels = unique(method)),
    yVal = as.numeric(yVal),
    predictVal = as.numeric(predictVal)
  )

ggplot(ctn_plot_data, aes(x = predictVal, y = yVal, color = method)) +
  geom_point(size = 0.5, alpha = 0.8) +         
  geom_abline(slope = 1, intercept = 0, alpha = 0.2) +  
  coord_fixed() +                                
  labs(
    y = "Testing Data Outcome",
    x = "Predicted Outcome on Testing Data"
  ) +                                            
  scale_x_continuous(breaks = NULL) +             
  scale_y_continuous(breaks = NULL) +           
  theme_bw() +                                  
  theme(
    axis.text = element_blank(),               
    strip.placement = "outside",               
    strip.background = element_blank()           
  ) +
  facet_grid(Y ~ method)                          

```

## Revised Stacking Algorithm for Binary Outcome

The MTPS package can fit models and make predictions when the
outcome is binary. We use the HIV data to illustrate the usage. First we
need convert the continuous outcome in the HIV data into binary outcome.
the HIV database website recommended cutoff values to convert IC50
ratios to binary outcomes are 3TC = 3, ABC = 2, AZT = 3, D4T = 1.5, and
DDI = 1.5. Then, using similar approach the data is split into
training data and testing data.

```{r bin-data, echo = T, results = 'hide', eval=T}
# https://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/CUTOFFS/drug.cutoffs
# cutoff value to be used to define drug resistent
cutoffs <- c(2,3,3,1.5,1.5) 
ymat_bin <- ymat
xmat_bin <- xmat
for(ii in 1:5) ymat_bin[,ii] <- (10^ymat[,ii] < cutoffs[ii]) * 1
y_train_bin <- ymat_bin[training_id, ]
y_test_bin  <- ymat_bin[-training_id, ]
x_train_bin <- xmat_bin[training_id, ]
x_test_bin  <- xmat_bin[-training_id, ]
```

For binary outcomes, the residual type can be `deviance', `pearson'
or `raw' for Deviance residual, Pearson residual or raw residual,
respectively. The default value for the ` r residual` is "deviance". The ` r resid.std` argument determines whether or
not to standardized the residuals and the default value is "FALSE".

Note that for Residual Stacking algorithm, the step 2 method must be
able to fit continuous outcomes. Though we are making predictions on
binary data, the step 2 in residual stacking is to make predictions on
the residuals Therefore, we need use a method that can fit models for
continuous outcomes.



The following code fit models using the Residual Stacking algorithms
with standardized Pearson residual. The step one uses tree method and
step two uses linear regression method.

```{r bin-train, echo = T, results = 'hide', eval= T}
fit_prs_std <- MTPS(xmat = x_train_bin, ymat = y_train_bin,
                               family = "binomial",
                               cv = FALSE, residual = TRUE,
                               method_step1 = rpart1,
                               method_step2 = lm1,
                               resid_type = "pearson", resid_std = TRUE) 
pred_prs_std <- predict(fit_prs_std, x_test_bin)
```

To see how the model performs we print the confusion matrices of each
outcome on the testing data. 0.5 is used as the threshold for the
prediction.

```{r bin-outcome, echo = F}
walk(1:ncol(y_test_bin), function(yy) {
  col_name <- colnames(y_test_bin)[yy]
  print(col_name)
  cm <- table((pred_prs_std[, yy] > 0.5) * 1, y_test_bin[, yy])
  print(cm)
})

```


## Revised Stacking Algorithm for Mix Outcome

For illustration purpose, the first three columns of the outcome in the
HIV data and last two columns of the binary outcome HIV data are
combined for the mix outcome example.

```{r mix-data, echo = T, results = 'hide', eval=T}
ymat.mix <- bind_cols(as_tibble(ymat)[, 1:3], as_tibble(ymat_bin)[, 4:5]) %>% as.matrix()
xmat.mix <- xmat
y_train_mix <- ymat.mix[training_id, ]
y_test_mix <- ymat.mix[-training_id, ]
x_train_mix <- xmat.mix[training_id, ]
x_test_mix <- xmat.mix[-training_id, ]
```

When the outcome variable is a combination of continuous and binary
outcomes, it is necessary to specify each outcome type in the ` r family` argument. For example, we fit the model using
Residual Stacking algorithm. The first step uses LASSO and the second
step uses tree method.

```{r mix-training, echo = T, results = 'hide', warning=FALSE, message=FALSE, eval=T}
fit_mix_rs <- MTPS(
  xmat = x_train_mix,
  ymat = y_train_mix,
  family = c("gaussian", "gaussian", "gaussian", "binomial", "binomial"),
  cv = FALSE,
  residual = TRUE,
  method_step1 = glmnet_lasso,
  method_step2 = rpart1
)
pred_mix_rs <- predict(fit_mix_rs, x_test_mix)
```

The following code draw the scatter plot of predicted value versus the actual outcomes on the first three columns of testing data, and the two confusion matrices for the last two columns of the testing data using 0.5 as the cutoff value. 

```{r mix-outcome, eval=T}
n_test <- nrow(x_test)

mix_plot_data <- tibble(
  Y = rep(colnames(y_test_mix)[1:3], each = nrow(y_test_mix)),
  predict = c(pred_mix_rs[, 1], pred_mix_rs[, 2], pred_mix_rs[, 3]),
  outcome = c(y_test_mix[, 1], y_test_mix[, 2], y_test_mix[, 3])
) %>%
  mutate(
    predict = as.numeric(predict),
    outcome = as.numeric(outcome)
  )

ggplot(mix_plot_data) +
  geom_point(aes(x = predict, y = outcome, color = Y),
             size = 0.5,
             alpha = 0.8) +
  ylab("Outcome of Testing Data") + xlab("Predicted Outcome of Testing Data") +
  scale_x_discrete(breaks = NULL) +
  scale_y_discrete(breaks = NULL) +
  geom_abline(slope = 1, alpha = 0.2) +
  coord_fixed() +
  theme_bw() +
  theme(
    legend.title = element_blank(),
    axis.text = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank()
  ) +
  facet_grid( ~ Y)
```
```{r eval=FALSE, include=FALSE}
ggplot(mix_plot_data, aes(x = predict, y = outcome, color = Y)) +
  geom_point(size = 0.5, alpha = 0.8) +
  geom_abline(slope = 1,
              intercept = 0,
              alpha = 0.2) +
  coord_fixed() +
  labs(y = "Outcome of Testing Data", x = "Predicted Outcome of Testing Data") +
  scale_x_continuous(breaks = NULL) +
  scale_y_continuous(breaks = NULL) +
  theme_bw() +
  theme(
    legend.title = element_blank(),
    axis.text = element_blank(),
    strip.placement = "outside",
    strip.background = element_blank()
  ) + facet_grid( ~ Y)
```

```{r echo=T, eval=T}
walk(4:5, function(yy) {
  col_name <- colnames(y_test_mix)[yy]
  print(col_name)
  cm <- table((pred_mix_rs[, yy] > 0.5) * 1, y_test_mix[, yy])
  print(cm)
})
```



## Residual Stacking for Survival Outcome

We simulated a survival dataset, simdat_mtps, to illustrate its usage. This dataset is designed for predicting multiple survival outcomes. The simulated data includes two outcome variables, and users can select multiple survival outcomes based on their needs.
\code{ymat} contains 2 survival response variables with 10\% censoring for 100 observations and \code{xmat} contains 2 predictors of those observations. The simulated survival time and status are used as multivariate outcomes, including \strong{time01},\strong{status01},\strong{time02},\strong{status02}. The predictors both follow a standard normal distribution. 

```{r data2}
data("simdat_mtps")
head(ymat)
head(xmat)

dim(ymat)
dim(xmat)
```

Currently, we provide two base learners to handle survival data:  ` r surv` and  ` r ela1`. When selecting the parametric AFT model (` r surv`) as the first-step learner, the following example can be executed:

```{r fit, echo = T, results = 'hide', warning=FALSE, message=FALSE, eval=FALSE}
# fit Residual Stacking Model for Survival Data
fit <- MTPS(
  xmat_train,
  ymat_train,
  xmat_list = xmat_list,
  family = 'survival',
  cv = FALSE,
  residual
  = TRUE,
  nfold = 5,
  method_step1 = surv,
  method_step2 = lm1,
  dist1 = "lognormal"
)
# predict the survival time on test set
pre3 <- predict.MTPS(fit, xmat_test)
```

When selecting the Elastic Net AFT model (` r ela1`) as the first-step learner, it is necessary to compute Kaplan-Meier estimators

```{r Simulat survival data, echo = T, results = 'hide', warning=FALSE, message=FALSE, eval=FALSE}
set.seed(1)
data("simdat_mtps")
# prepare training and test set
id.train <- sample(1:100, 80)
xmat_train <- xmat[id.train, ]
xmat_test <- xmat[-id.train, ]
ymat_train <- cbind(list(survival::Surv(ymat[id.train, "time01"], ymat[id.train, "status01"])), list(survival::Surv(ymat[id.train, "time02"], ymat[id.train, "status02"])))

# Produce the Kaplan-Meier estimator
weights <- find_km_weights_mat(ymat[id.train, ], num_outcome = 2)

# fit Residual Stacking Model for Survival Data
fit <- MTPS(
  xmat_train,
  ymat_train,
  xmat_list = xmat_list,
  family = 'survival',
  cv = FALSE,
  residual
  = TRUE,
  nfold = 5,
  method_step1 = ela1,
  method_step2 = lm1,
  dist1 = "lognormal"
)

# predict the survival time on test set
pre4 <- predict.MTPS(fit, xmat_test)
```



# Specifying Models in Step 1 and Step 2

## Specify Different Base Learners for Different Outcomes within One Step

In the above examples, only one base learner is used in each step. The
MTPS package allow users to choose different base learners
within one step for each outcome according to the outcome properties.
The following example use different base learners for different
outcomes. In step one, the first three outcomes are fitted by the LASSO
model and the last two outcomes are fitted by the linear regression
methods. In the second step, the first outcome is fitted by the tree
methods and the last four outcome is fitted by the LASSO model. By
default the residual stacking algorithm is applied.

```{r mix-mtd, echo = T, warning=FALSE,message=FALSE, eval=FALSE}
fit_mixOut <- MTPS(
  xmat = x_train,
  ymat = y_train,
  family = "gaussian",
  method_step1 =
    c(glmnet_lasso, glmnet_lasso, glmnet_lasso, lm1, lm1),
  method_step2 =
    c(rpart1, glmnet_lasso, glmnet_lasso, glmnet_lasso, glmnet_lasso)
)

pred <- predict(fit_mixOut, x_test)
```

## User-Defined Base Learners

Apart from giving base learners in the package, users are able to define
new methods to fit models in step one or step two. Users can either change
default parameters in our provided base learners or define their new
methods.

### Modify Arguments for Existing Base Learners
The MTPS package provides several methods to fit models. Some
of these methods has some default arguments. For example, the glmnet1
method fit a generalized linear model with elasticnet regularization. By
default, using cross validation, the model chose the best alpha in 0,
0.2, 0.4, 0.6, 0.8 and 1. If user want to fit a LASSO or ridge model,
that is, fix the value of alpha equals 1 or 0. (glmnet_lasso and
glmnet_ridge method are available in the MTPS package) users
can use the ` r modify.parameter` function to pass arguments.

```{r method-mod, eval=F, echo=T, eval=T}
glmnet_lasso <- modify_parameter (glmnet1, alpha = 1)
glmnet_ridge <- modify_parameter (glmnet1, alpha = 0)
```

### Define New Base Learners

Apart from modifying model parameters, users can define new methods. In
order to use the new methods, users need to make sure that the input and
output format is consistent with existing methods.

The new method should be defined as a function with predictor xmat
(matrix), outcome ymat (matrix with 1 column or a vector), and the
outcome family as inputs. The outputs of the function should be a list
which contains three elements called "model", "y.fitted" and "predFun". The "model" element is the method object (which will be used in the corresponding predict function). The "y.fitted" element should be a matrix / dataframe of fitted value and the "predFun" should be a ` r predict` function for the method object. 

The following example shows how to define a method.

```{r method-new, eval=F, echo=T,eval=FALSE}
glm1 <- function(xmat, ymat, family, ...) {
  tmp0 <- data.frame(yy = ymat, xmat)
  model <- glm(yy ~ ., data = tmp0, family = family, ...)
  y.fitted <- fitted(model)
  predFun <- function(model, xnew) {
    predict(model, newdata = data.frame(xnew), type = "response")
  }
  return(list(
    model = model,
    y.fitted = y.fitted,
    predFun = predFun
  ))
}
```

After defining new base learners, Users can pass these methods in method
arguments in the MTPS function. 

# Performance Comparison

Cross-Validation is a popular way to compare different methods. Here the
cross-validation method is used as an example to evaluate the
performance of choice of base learners in the stacking algorithms. The ` r cv.MTPS` function provide a simple interface for
model evaluation using cross-validation. It returns the mean square
error for continuous outcomes and AUC, accuracy, recall and precision for
binary outcomes from cross-validation predictions.

Following is an example to use 5-fold Cross-Validation method to compare
using LASSO and ridge models in step 1 of residual stacking. Both of
them are followed by linear regression in step 2.

The cross-validation process can be repeated to compare the these two
methods. For speed reasons here we only repeat the above process for 20
times and plot the boxplot of their MSE.

```{r learner-compare, echo=T,eval=T}
nsim <- 20
mse_lasso_lm <- matrix(NA, nrow = nsim, ncol = ncol(ymat)) %>%
  as_tibble(.name_repair = "minimal") %>%
  set_names(colnames(ymat))
mse_ridge_lm <- mse_lasso_lm

for (ii in 1:nsim) {
  set.seed(ii)
  # lasso stacking with lm
  mse_lasso_lm[ii,] <- cv_MTPS(xmat, ymat, family="gaussian",
                            cv = FALSE, residual = TRUE,
                            method_step1 = glmnet_lasso, method_step2 = lm1,
                            resid_std = FALSE)$continuous
  # ridge stacking with lm
  mse_ridge_lm[ii,] <- cv_MTPS(xmat, ymat, family="gaussian",
                            cv = FALSE, residual = TRUE,
                            method_step1 = glmnet_ridge, method_step2 = lm1,
                            resid_std = FALSE)$continuous
}

mse_data <- tibble(
  lasso = rowMeans(mse_lasso_lm, na.rm = TRUE),
  ridge = rowMeans(mse_ridge_lm, na.rm = TRUE)
)

mse_data <- mse_data %>%
  pivot_longer(cols = everything(), names_to = "Learner", values_to = "MSE")

ggplot(mse_data) +
  geom_boxplot(aes(x=Learner, y=MSE, fill=Learner)) +
  ggtitle("Boxplot of average Mean Square Error") +
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_blank())
```
```{r eval=FALSE, include=FALSE}

ggplot(mse_data, aes(x = Learner, y = MSE, fill = Learner)) +
  geom_boxplot() +
  ggtitle("Boxplot of Average Mean Square Error") +
  theme_bw() +
  theme(
    legend.position = "none",  
    plot.title = element_text(hjust = 0.5),  
    axis.title.x = element_blank()  
  ) +
  scale_fill_manual(values = c("lasso" = "#1f78b4", "ridge" = "#33a02c")) 

```


Similarly, the above cross-validation process can be applied when the
outcome is binary or a combination of continuous and binary outcomes.
Some possible evaluation methods are accuracy rate, area under curve
(AUC), etc.

When dealing with the survival outcomes, we use the \code{mse()} and \code{c_index()} to calculate the MSE and C-index of the predicted outcomes without considering censor observations. Note that the MSE here is computed on the log-scale.
```{r fomula of MSE and C-index,echo = T, results = 'hide', warning=FALSE, message=FALSE, eval=FALSE}

# MSE 
mse <- function(pre, object){
  # pre_time: the predicted event time
  # obj: 2 columns, time & status
  ind <- which(object$status == 1)
  event_time <- object$time[ind]
  pre_event_time <- pre[ind]
  names(event_time) <- names(pre_event_time)
  return (mean((log(event_time) - log(pre_event_time))^2))
}

# C-index
c_index <- function(pre, object){
  total.pairs <- 0
  c <- 0
  test <- cbind(pre, object)
  test <- test[order(test[,2]),]
  for (i in 1:nrow(test)){
    if(i ==nrow(test)) break
    else if(test[i,3]==0) next # only consider uncensored observations
    else{
      for (j in (i+1):nrow(test)){
        total.pairs <- total.pairs+1
        if(test[j,2] == test[i,2]) {if(test[j,1] > test[i,1] & test[j,1] == test[i,1]) c <- c + 0.5}
        #"if(test[j,1] > test[i,1] & test[j,1] == test[i,1]) c <- c + 0.5"
        #if we want add all possible results we can use straightly "c <- c + 0.5"
        
        if(test[j,2] > test[i,2]){if(test[j,1] > test[i,1]) {c <- c+1}}
        if(test[j,2] > test[i,2]){if(test[j,1] == test[i,1]) {c <- c+0.5}}
      }
    }
    
  }
  return (c/total.pairs)
}


```

# Reference


<p>Dimitriadou, E., Hornik, K., Leisch, F., Meyer, D., and Weingessel, A. (2008). <br>Misc functions of the Department of Statistics (e1071), TU Wien.<br>
<em>R package</em>,1,5-24.</p>

<p>Friedman, J., Hastie, T., and Tibshirani, R. (2010).<br>Regularization Paths for Generalized Linear Models via Coordinate Descent<br>
<em>Journal of Statistical Software</em>, Vol. 33(1) 1-22.</p>

<p>Kuhn, M. (2008).<br>Building Predictive Models in R Using the caret Package.<br>
<em>Journal of Statistical Software</em>, Vol. 28(5), 1-26.</p>

<p>Ripley, B.D. (1996).<br>Pattern Recognition and neural networks<br>
<em>Cambridge University Press.</em></p>

<p>Venables, W. N., & Ripley, B. D. (2013).
<em>Modern applied statistics with S-PLUS.</em>Springer Science & Business Media.</p>

<p>Xing, L., Lesperance, M., \& Zhang, X. (2019).<br>Simultaneous prediction of multiple outcomes using revised stacking algorithms<br>
<em>Bioinformatics</em>, https://doi.org/10.1093/bioinformatics/btz531.</p>



